## Young-of-year size model {#modelYOY}

```{r globalModelsYOY, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r librariesModelsYOY, echo = FALSE}
library(getWBData)
library(tidyverse)
library(lubridate)
library(kableExtra)

```

The question here is what is driving _body size variation_ across years in brook trout and brown trout in the WB?  

We focus on ageInSamples == 1 (age-0 fish in the fall sample) fish for growth model. This is the first sampling occasion that most fish are big enough to tag. Not all fish are big enough, however, and there is a number of untagged fish each year. We need to include both tagged and untagged fish in our age-0 size model.

Factors to include in the model are   
1. Sample date  
2. Cumulative temperature prior to sampling  
3. Cumulative flow prior to sampling  
4. Extreme flow events?? Floods, droughts?  
5. Fish density, age-0 counts across all three salmonids  

### Raw data for YOY model
1. Environmental data (flow, temperature) are from 1, 3, or 5 months prior to date of individual capture. Also can used fixed dates: assumed spawning dates, assumed emergence dates and actual observation (sample) dates.
2. All fish data are from age-0 in autumn.
3. Abundance data.

### Get environmental data

West Brook environmental data (flow and temperature)
```{r envDataWB, cache = TRUE}
load(file = './data/envDataWB.RData')

```


### Get first observations

1. Filter cdWB_electro for first observations in the autumn for age-0 fish (ageInsamples == 1). Including both tagged and untagged fish.
```{r firstObs0}
selectedVariables <- c("tag", "species", "river", "detectionDate", "sampleNumber", "n", "proportionSampled", "observedLength", "observedWeight", "area", "section", "season", "isYOY")

firstObs_noTag <- cdWB_electro %>%
  filter(is.na(tag), ageInSamples == 1) %>%
  mutate(n = 1) %>%
  dplyr:: select(all_of(selectedVariables)) 

firstObs_tag <- cdWB_electro %>%
  group_by(tag) %>%
  mutate(isFirstObs = detectionDate == min(detectionDate),
         n = n()) %>%
  filter(isFirstObs, ageInSamples == 1) %>%
  dplyr::select(all_of(selectedVariables)) %>%
  ungroup()

firstObs0 <- add_row(firstObs_tag, firstObs_noTag) %>%
  mutate(date = as_date(detectionDate),
         yday = yday(date),
         year = year(date))

```


For each date in firstObs0 that at least one fish was captured, calculate summary stats for flow and temperature for different time periods:   
1. Assumed spawning to capture  
2. Assumed spawning to assumed emergence  
3. Assumed emergence to capture  
4. One month preceding capture  
5. Three months preceding capture  
5. Five months preceding capture


Then merge results with firstObs0 to create firstObs.
```{r firstObs}

spawn_month <- "11" # spawning
spawn_day <- "15"
emerge_month <- "03" # emergence
emerge_day <- "01"

firstObsDates <- firstObs0 %>% distinct(date = date(detectionDate), river)

# move to getPrepareWBData
getEnvMeans <- function(riverIn, start, end) { 
  out <- envDataWB %>% 
    filter(river == riverIn, dateDate >= start, dateDate <= end) %>%
    summarize(
      sumT = sum(temperature, na.rm = TRUE),
      meanT = mean(temperature, na.rm = TRUE),
      sdT = sd(temperature, na.rm = TRUE), 
      cvT = sdT/meanT,
      
      sumF = sum(flow, na.rm = TRUE),
      meanF = mean(flow, na.rm = TRUE),
      sdF = sd(flow, na.rm = TRUE),
      cvF = sdF/meanF,
      n = n()
    )

  #message(paste(river, start, end,tag))
  return(out)
}

firstObs_Env <- firstObsDates %>%
  rowwise() %>%
  mutate(
    year = year(date),
    spawnDate = ymd(paste0(year,spawn_month,spawn_day)) - years(1),
    emergeDate = ymd(paste0(year,emerge_month,emerge_day)),
    oneMonthDate = date - days(as.integer(1 * 30.5)), #months(1), 'months gives error when prev month has 30 days and current has 31
    threeMonthDate = date - days(as.integer(3 * 30.5)),
    fiveMonthDate = date - days(as.integer(5 * 30.5)),
    spawn_emerge = list(getEnvMeans(river, spawnDate, emergeDate)),
    emerge_detect = list(getEnvMeans(river, emergeDate, date)),
    spawn_detect = list(getEnvMeans(river, spawnDate, date)),
    oneMonth = list(getEnvMeans(river, oneMonthDate, date)),
    threeMonth = list(getEnvMeans(river, threeMonthDate, date)),
    fiveMonth = list(getEnvMeans(river, fiveMonthDate, date))
  )

# merge env data into firstObs0
firstObs <- firstObs0 %>%
  left_join(firstObs_Env)

#str(firstObs)
```

Unnest firstObs so environmental summary stats are available as data frame with the name of the time interval as the prefix to the statisticVariable name
```{r firstObsUnnested}

getScaled <- function(d){
  (d - mean(d, na.rm = TRUE)) / sd(d, na.rm = TRUE)
}

# this scales across all individuals - I think this is ok
firstObsUnnested <- firstObs %>% 
  unnest(cols = c(spawn_emerge, emerge_detect, spawn_detect, oneMonth, threeMonth, fiveMonth), names_sep = "_") %>%
  mutate(
    emerge_detect_sumTScaled = getScaled(emerge_detect_sumT),
    emerge_detect_sumFScaled = getScaled(emerge_detect_sumF),
    oneMonth_sumTScaled = getScaled(oneMonth_sumT),
    oneMonth_sumFScaled = getScaled(oneMonth_sumF),
    threeMonth_sumTScaled = getScaled(threeMonth_sumT),
    threeMonth_sumFScaled = getScaled(threeMonth_sumF),
    fiveMonth_sumTScaled = getScaled(fiveMonth_sumT),
    fiveMonth_sumFScaled = getScaled(fiveMonth_sumF),
    ydayScaled = getScaled(yday)
  )

str(firstObsUnnested)
#ggplot(firstObsUnnested, aes(oneMonth_sumTScaled, fiveMonth_sumTScaled)) +
#  geom_point() +
#  facet_wrap(~river)
```

### Counts of captured fish
Min and max years (inclusive) for standardizing counts
```{r minMaxYear}
minYear <- 2000
maxYear <- 2015
```

Counts by river and species
```{r rawCounts}

countsRSY <- firstObs %>%
  filter(year %in% minYear:maxYear) %>%
  group_by(river, species, year) %>%
  summarize(
    count = n(),
    meanPropSampled = mean(proportionSampled, na.rm = TRUE)
  ) %>%
  mutate(countAdj = count / meanPropSampled) %>%
  ungroup() %>%
  group_by(river, species) %>%
  mutate(meanCountRS = mean(count, na.rm = TRUE),
         sdCountRS = sd(count, na.rm = TRUE),
         countRS_Scaled = (count - meanCountRS) / sdCountRS) %>%
  ungroup()

ggplot(countsRSY, aes(year, countRS_Scaled, color = species)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ river)
```

Counts by river
```{r countsRY}
countsRY <- firstObs %>%
  filter(year %in% minYear:maxYear) %>%
  group_by(river, year) %>%
  summarize(
    count = n(),
    meanPropSampled = mean(proportionSampled, na.rm = TRUE)
  ) %>%
  mutate(countAdj = count / meanPropSampled) %>%
  ungroup() %>%
  group_by(river) %>%
  mutate(meanCountR = mean(count, na.rm = TRUE),
         sdCountR = sd(count, na.rm = TRUE),
         countR_Scaled = (count - meanCountR) / sdCountR) %>%
  ungroup()

ggplot(countsRY, aes(year, countR_Scaled, color = river)) +
  geom_point() +
  geom_line()
```

Counts for the metaPopulation (WB, Jimmy, Mitchell)  
Use these for modelling.
```{r countsMeta}
countsMetaY <- firstObs %>%
  filter(river != "wb obear", year %in% minYear:maxYear) %>%
  group_by(year) %>%
  summarize(
    count = n(),
    meanPropSampled = mean(proportionSampled, na.rm = TRUE)
  ) %>%
  mutate(countAdj = count / meanPropSampled) %>%
  ungroup() %>%
  mutate(meanCount = mean(count, na.rm = TRUE),
         sdCount = sd(count, na.rm = TRUE),
         count_Scaled = (count - meanCount) / sdCount)
# missing data for tribs in 2000, 2001 - may skew scaled count a bit low - should fix

ggplot(countsMetaY, aes(year, count_Scaled)) +
  geom_point() +
  geom_line()
```

Merge metapopulation scaled counts into firstObsUnnested
```{r mergeMeta}

firstObsUnnested <- firstObsUnnested %>%
  left_join(countsMetaY %>% dplyr::select(year, count_Scaled))

firstObsUnnestedWB <- firstObsUnnested %>% filter(river == "west brook")
```


### Raw data plots

#### Frequency plots by species and river

#### Brook Trout, West brook  
```{r raw data plots1}
  #cd1 <- cdWB_electro %>% filter(ageInSamples == 1, species != 'ats')

  plotSppRiv = function(s, r) { 
    ggplot(firstObs %>% filter(species == s, river == r), aes(observedLength, color = is.na(tag))) +
      geom_freqpoly() +
      geom_vline(xintercept = 60) +
      ggtitle(paste(s, r, sep = ', ')) +
      xlim(c(30,125)) +
      facet_wrap(~ year, scales = "free_y")
  }
  
  species = 'bkt'
  riverOrdered = "west brook"

  plotSppRiv(species, riverOrdered)
```  

#### Brook Trout, wb jimmy 
```{r raw data plots2, echo = FALSE}
  species = 'bkt'
  riverOrdered = "wb jimmy"

  plotSppRiv(species, riverOrdered)
```  

#### Brook Trout, wb mitchell  
```{r raw data plots3, echo = FALSE}
  species = 'bkt'
  riverOrdered = "wb mitchell"

  plotSppRiv(species, riverOrdered)
```  

#### Brook Trout, wb obear  
```{r raw data plots4, echo = FALSE}
  species = 'bkt'
  riverOrdered = "wb obear"

  plotSppRiv(species, riverOrdered)
```  

#### Brown Trout, West brook  
```{r raw data plots5, echo = FALSE}
  species = 'bnt'
  riverOrdered = "west brook"

  plotSppRiv(species, riverOrdered)
```  

#### Brown Trout, wb jimmy 
```{r raw data plots6, echo = FALSE}
  species = 'bnt'
  riverOrdered = "wb jimmy"

  plotSppRiv(species, riverOrdered)
```  

#### Brown Trout, wb mitchell  
```{r raw data plots7, echo = FALSE}
  species = 'bnt'
  riverOrdered = "wb mitchell"

  plotSppRiv(species, riverOrdered)
```  

#### Brown Trout, wb obear - there are no Brown trout in O'Bear  

#### Trout, in the WB mainstem only
```{r troutWB}
  
  ggplot(firstObs %>% filter(species != "ats"), aes(observedLength)) +
    geom_freqpoly() +
    geom_vline(xintercept = 60, color = 'orange') +
    facet_grid(species ~ year)
  
  ggplot(firstObs %>% filter(species != "ats"), aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60, color = 'orange') +
    facet_grid(species ~ year)
```
  

#### Why are there untagged fish bigger than 60mm?  
Check 2002/bkt/WB, as an example
Answer: because they are outside the study area (area = 'above' or 'below') or were tagging mortalities

```{r 2002 untagged}

  firstObs2002BKT <- firstObs %>% filter(year == 2002, species == "bkt")
  table(is.na(firstObs2002BKT$tag))
  
  ggplot(firstObs2002BKT, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60)
  
  # looks like untagged area=inside fish wee morts, the rest were above or below
  ggplot(firstObs2002BKT, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60) +
    facet_grid(~area)
  
  # check 2003
  ggplot(firstObs %>% filter(year == 2003, species == "bkt"), aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60) +
    facet_grid(~area)
  # same story as 2002
```

#### Why no untagged fish for 2000 and 2001?
Check data logs to see if we were not recording untagged fish in 2000, 2001
```{r untagged 2000, 2001}

  cfirstObs2000_2001BKT <- firstObs %>% filter(year %in% 2000:2001, species == "bkt")
  table(is.na(cfirstObs2000_2001BKT$tag))
  table(cfirstObs2000_2001BKT$observedLength)

  cfirstObs2000_2001BKT %>% filter(observedLength < 60)
  
  ggplot(cfirstObs2000_2001BKT, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60)
```


### Models based on yearly means

Filter firstObsUnnestedWB for bkt, bnt and min/maxYear
```{r filter firstObsUnnestedWB}
d_WB_BKT_BNT <- firstObsUnnestedWB %>% filter(species != "ats", year %in% minYear:maxYear) %>%
  mutate(species01 = ifelse(species == "bkt", 1, 0))

hist(d_WB_BKT_BNT$detectionDate, breaks = 250)

d_BKT_BNT <- firstObsUnnested %>% filter(species != "ats", year %in% minYear:maxYear) %>%
  mutate(species01 = ifelse(species == "bkt", 1, 0))

```

Mean model functions
```{r meanModelFunctions}

getMeansData <- function(d, t, f) {
  means <- d %>% 
    group_by(species, year) %>% 
    summarize(meanLength = mean(observedLength, na.rm = TRUE), 
              meanEmerge_detect_sumTScaled = mean(emerge_detect_sumTScaled, na.rm = TRUE),
              meanEmerge_detect_sumFScaled = mean(emerge_detect_sumTScaled, na.rm = TRUE),
              meanTTime_sumTScaled = mean(get(t), na.rm = TRUE),
              meanFTime_sumFScaled = mean(get(f), na.rm = TRUE),
              meanYdayScaled = mean(ydayScaled, na.rm = TRUE),
              meanCount_Scaled = mean(count_Scaled, na.rm = TRUE)
              )
  return(means)
}

getMeansDataByRiver <- function(d, t, f) {
  means <- d %>% 
    group_by(species, year, river) %>% 
    summarize(meanLength = mean(observedLength, na.rm = TRUE), 
              meanEmerge_detect_sumTScaled = mean(emerge_detect_sumTScaled, na.rm = TRUE),
              meanEmerge_detect_sumFScaled = mean(emerge_detect_sumTScaled, na.rm = TRUE),
              meanTTime_sumTScaled = mean(get(t), na.rm = TRUE),
              meanFTime_sumFScaled = mean(get(f), na.rm = TRUE),
              meanYdayScaled = mean(ydayScaled, na.rm = TRUE),
              meanCount_Scaled = mean(count_Scaled, na.rm = TRUE)
              )
  return(means)
}

plotMeans <- function(means){
  out <- list()
  out[[1]] <- ggplot(means, aes(meanTTime_sumTScaled, meanLength, color = species)) +
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
  
  out[[2]] <- ggplot(means, aes(meanFTime_sumFScaled, meanLength, color = species)) +
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
  
  out[[3]] <- ggplot(means, aes(meanTTime_sumTScaled, meanFTime_sumFScaled, color = species)) +
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
  return(out)
}

runMeanModels <- function(means) {
  modLMMeans1 <- lm(meanLength ~ (factor(species) + meanFTime_sumFScaled + meanTTime_sumTScaled + meanYdayScaled + meanCount_Scaled), data = means)
  modLMMeans2 <- lm(meanLength ~ (factor(species) + meanFTime_sumFScaled + meanTTime_sumTScaled + meanYdayScaled + meanCount_Scaled)^2, data = means)
  modLMMeans3 <- lm(meanLength ~ (factor(species) + meanFTime_sumFScaled + meanTTime_sumTScaled + meanYdayScaled + meanCount_Scaled)^3, data = means)
  return(list(modLMMeans1, modLMMeans2, modLMMeans3))
}
```

Mean lengths by river. This is information only. Using the WB data only shown here and in the next graph for the models.
```{r meanLengthsR, echo = FALSE}
means1R <- getMeansDataByRiver(d_BKT_BNT, "oneMonth_sumTScaled", "oneMonth_sumFScaled")
  ggplot(means1R, aes(year, meanLength, shape = species, color = species)) + 
    geom_point() +
    geom_line() +
    facet_wrap(~river)
```

Mean lengths for the mean length model.
```{r meanLengths, echo = FALSE}    
means1 <- getMeansData(d_WB_BKT_BNT, "oneMonth_sumTScaled", "oneMonth_sumFScaled")
  ggplot(means1, aes(year, meanLength, color = species)) + 
    geom_point() +
    geom_line()
```

Graphs for variables that do not depend on number of months
```{r staticGraphs, echo = FALSE}
  ggplot(means1, aes(meanYdayScaled, meanLength, color = species)) +
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
  
  ggplot(means1, aes(meanCount_Scaled, meanLength, color = species)) +
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE)
  
```


#### Models with flow and temperature from previous *one* month
```{r means1Month, echo = FALSE}
#means1 calculated above
plotMeans(means1)
mod1 <- runMeanModels(means1)

AIC(mod1[[1]], mod1[[2]], mod1[[3]]) %>% arrange(AIC)

summary(mod1[[1]])
rI1 <- relaimpo::calc.relimp(mod1[[1]])

save(means1, mod1, file = "models/mod1LMMeans.Rdata")
```

Relative importance for main effects model
```{r rI1, echo = FALSE}
rI1$lmg
```


#### Models with flow and temperature from previous *three* months
```{r means3Month, echo = FALSE}
means3 <- getMeansData(d_WB_BKT_BNT, "threeMonth_sumTScaled", "threeMonth_sumFScaled")
plotMeans(means3)
mod3 <- runMeanModels(means3)

AIC(mod3[[1]], mod3[[2]], mod3[[3]]) %>% arrange(AIC)

summary(mod3[[1]])

rI3 <- relaimpo::calc.relimp(mod3[[1]])

save(means3, mod3, file = "models/mod3LMMeans.Rdata")
```

Relative importance for main effects model
```{r rI3, echo = FALSE}
rI3$lmg
```


#### Models with flow and temperature from previous *five* months
```{r means5Month, echo = FALSE}
means5 <- getMeansData(d_WB_BKT_BNT, "fiveMonth_sumTScaled", "fiveMonth_sumFScaled")
plotMeans(means5)
mod5 <- runMeanModels(means5)

AIC(mod5[[1]], mod5[[2]], mod5[[3]]) %>% arrange(AIC)

summary(mod5[[1]])

rI5 <- relaimpo::calc.relimp(mod5[[1]])

save(means5, mod5, file = "models/mod5LMMeans.Rdata")
```

Relative importance for main effects model
```{r rI5, echo = FALSE}
rI5$lmg
```


r-squared values and AICs for 1st, 2nd (2-way interactions) and 3rd (3-way interactions) order models
```{r r2, echo = FALSE}
knitr::kable(
 tibble(Order = c(1,2,3),r2 = c(summary(mod1[[1]])$r.squared, summary(mod1[[2]])$r.squared, summary(mod1[[3]])$r.squared)) %>% add_column(numMonths = 1),
 digits = 3
) %>%
  kable_styling(full_width = FALSE)

knitr::kable(
 tibble(Order = c(1,2,3),r2 = c(summary(mod3[[1]])$r.squared, summary(mod3[[2]])$r.squared, summary(mod3[[3]])$r.squared)) %>% add_column(numMonths = 3),
 digits = 3
) %>%
  kable_styling(full_width = FALSE)

knitr::kable(
 tibble(Order = c(1,2,3),r2 = c(summary(mod5[[1]])$r.squared, summary(mod5[[2]])$r.squared, summary(mod5[[3]])$r.squared)) %>% add_column(numMonths = 5),
 digits = 3
) %>%
  kable_styling(full_width = FALSE)

# AIC
knitr::kable(
  format = 'html',
  AIC(mod1[[1]], mod1[[2]], mod1[[3]]) %>% arrange(AIC) %>% add_column(numMonths = 1),
  digits = 3
) %>%
  kable_styling(full_width = FALSE)

knitr::kable(
  AIC(mod3[[1]], mod3[[2]], mod3[[3]]) %>% arrange(AIC) %>% add_column(numMonths = 3),
  digits = 3
) %>%
  kable_styling(full_width = FALSE)

knitr::kable(
  AIC(mod5[[1]], mod5[[2]], mod5[[3]]) %>% arrange(AIC) %>% add_column(numMonths = 5),
  digits = 3
) %>%
  kable_styling(full_width = FALSE)

```

Relative importance of main effects models (repeat of above, but all in one place here)
```{r rIKable, echo = FALSE}
library("kableExtra")

# Relative Importance
knitr::kable(
 tibble(var = names(rI1$lmg), relImp = rI1$lmg) %>% add_column(numMonths = 1),
 digits = 3
) %>%
  kable_styling(full_width = FALSE)

knitr::kable(
 tibble(var = names(rI3$lmg), relImp = rI3$lmg) %>% add_column(numMonths = 3),
 digits = 3
) %>%
  kable_styling(full_width = FALSE)

knitr::kable(
 tibble(var = names(rI5$lmg), relImp = rI5$lmg) %>% add_column(numMonths = 5),
 digits = 3
) %>%
  kable_styling(full_width = FALSE)
```

### Models with extreme flow events (droughts)
We get negative cumulFlows because we have some negative flows from the flow extension model
```{r drought}
# put some of these calculations into envDataWB
envDataWBFlow = envDataWB %>%
  mutate(year = year(dateDate),
         yday = yday(dateDate),
         flowNoNAs = ifelse(is.na(flow), 0, flow),
         tempNoNAs = ifelse(is.na(temperature), 0, temperature)) %>%
  filter(year %in% minYear:maxYear, 
         yday > 100, yday < 300,
         river == "west brook") %>%
  group_by(year) %>%
  mutate(cumulFlow = cumsum(flowNoNAs),
         cumulFlow01 = cumulFlow / max(cumulFlow),
         cumulTemp = cumsum(tempNoNAs)) %>%
  ungroup()

firstObsYears <- firstObs %>%
    filter(year %in% minYear:maxYear, 
           yday > 100, yday < 300)

ggplot(envDataWBFlow, aes(yday, flow)) +
  geom_point(aes(yday, observedLength/20), size = 0.75, alpha = 0.2, color = 'lightblue', data = firstObsYears) +
  geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(0,300, 30)) +
  facet_wrap(~year)

ggplot(envDataWBFlow, aes(yday, cumulFlow / 10)) +
  geom_point(aes(yday, observedLength / 20), size = 0.75, alpha = 0.2, color = 'lightblue', data = firstObsYears) +
  geom_point(size = 0.5, color = 'darkgrey') +
  geom_point(aes(yday, cumulTemp / 800), size = 0.5, color = "orange", data = envDataWBFlow) +
  geom_point(aes(yday, flow), size = 0.5, data = envDataWBFlow) +
  scale_x_continuous(breaks = seq(0, 300, 30)) +
  #theme_publication() +
  facet_wrap(~year)

ggplot(envDataWBFlow, aes(yday, cumulFlow, color = (year))) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 300, 30)) 

ggplot(envDataWBFlow, aes(yday, cumulTemp, color = (year))) +
  geom_point() +
  #geom_point(aes(yday, cumulTemp / 800, color = factor(year)), data = tmp) +
  scale_x_continuous(breaks = seq(0, 300, 30))

```

Is there a sampling section effect?

Note: there are fish in sections > 50 for years 2002 and 2003, need to filter out early
```{r}
ggplot(d_WB_BKT_BNT %>% filter( section <= 47), aes(factor(section), observedLength)) +
  geom_boxplot() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(d_WB_BKT_BNT %>% filter( section <= 47), aes(factor(year), observedLength)) +
  geom_boxplot() +
  geom_smooth() +
  facet_wrap(~section)
```

### Models based on individual observations
Probably not use these, too much individual variation


#### Are flow and temperature correlated for individual observations?
```{r flowTempCor, cache = TRUE}

ggplot(firstObsUnnestedWB, aes(oneMonth_sumTScaled, oneMonth_sumFScaled)) +
  geom_point(aes(color = (year))) +
  geom_smooth(method = "lm") +
  facet_wrap(~ species)

ggplot(firstObsUnnestedWB, aes(threeMonth_sumTScaled, threeMonth_sumFScaled)) +
  geom_point(aes(color = (year))) +
  geom_smooth(method = "lm") +
  facet_wrap(~ species)

ggplot(firstObsUnnestedWB, aes(fiveMonth_sumTScaled, fiveMonth_sumFScaled)) +
  geom_point(aes(color = (year))) +
  geom_smooth(method = "lm") +
  facet_wrap(~ species)

# water getting warmer during a sample
ggplot(firstObsUnnestedWB, aes(fiveMonth_sumTScaled, fiveMonth_sumFScaled)) +
  geom_point(aes(color = (yday))) +
  geom_smooth(method = "lm") +
  facet_wrap(~ species)

cor(firstObsUnnestedWB$oneMonth_sumTScaled, firstObsUnnestedWB$oneMonth_sumFScaled)
cor(firstObsUnnestedWB$threeMonth_sumTScaled, firstObsUnnestedWB$threeMonth_sumFScaled)
cor(firstObsUnnestedWB$fiveMonth_sumTScaled, firstObsUnnestedWB$fiveMonth_sumFScaled)
# brook trout
firstObsUnnestedWB_BKT <- firstObsUnnestedWB %>% filter(species == "bkt")
cor(firstObsUnnestedWB_BKT$oneMonth_sumTScaled, firstObsUnnestedWB_BKT$oneMonth_sumFScaled)
cor(firstObsUnnestedWB_BKT$threeMonth_sumTScaled, firstObsUnnestedWB_BKT$threeMonth_sumFScaled)
cor(firstObsUnnestedWB_BKT$fiveMonth_sumTScaled, firstObsUnnestedWB_BKT$fiveMonth_sumFScaled)

# brown trout
firstObsUnnestedWB_BNT <- firstObsUnnestedWB %>% filter(species == "bnt")
cor(firstObsUnnestedWB_BNT$oneMonth_sumTScaled, firstObsUnnestedWB_BNT$oneMonth_sumFScaled)
cor(firstObsUnnestedWB_BNT$threeMonth_sumTScaled, firstObsUnnestedWB_BNT$threeMonth_sumFScaled)
cor(firstObsUnnestedWB_BNT$fiveMonth_sumTScaled, firstObsUnnestedWB_BNT$fiveMonth_sumFScaled)
```

Do fish from long samples get bigger over time?  
No clear evidence.
```{r}
ggplot(firstObsUnnestedWB, aes(yday, observedLength)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm") +
  facet_wrap(~year)
```


Assign the month interval (one, three, five) for flow and temperature variables. The variable will be accessed using e.g. `get(tTime)` in formulas and filters
```{r time}
tTime <- "threeMonth_sumTScaled"
fTime <- "threeMonth_sumFScaled" #"fiveMonth_sumFScaled"

```


```{r rI_LM}
library(lme4)
library(relaimpo)


modLM1 <- lm(observedLength ~ (factor(species) * get(tTime) * get(fTime) * ydayScaled * count_Scaled), data = d_WB_BKT_BNT)

modLM2 <- lm(observedLength ~ (factor(species) + get(tTime) + get(fTime) + ydayScaled + count_Scaled), data = d_WB_BKT_BNT)

modLM3 <- lm(observedLength ~ (factor(species) + get(tTime) + get(fTime) + ydayScaled + count_Scaled)^2, data = d_WB_BKT_BNT)

modLM1a <- lm(observedLength ~ (factor(species)), data = d_WB_BKT_BNT)
modLM1b <- lm(observedLength ~ (factor(species) * get(tTime)), data = d_WB_BKT_BNT)
modLM1c <- lm(observedLength ~ (factor(species) * get(fTime)), data = d_WB_BKT_BNT)
modLM1d <- lm(observedLength ~ (factor(species) * ydayScaled), data = d_WB_BKT_BNT)
modLM1e <- lm(observedLength ~ (factor(species) * count_Scaled), data = d_WB_BKT_BNT)

AIC(modLM1, modLM2, modLM3, modLM1a, modLM1b, modLM1c, modLM1d, modLM1e) %>% arrange(AIC)

#relaimpo::calc.relimp(modLM2) # slow for bigger models


# get 'boundary (singular)' error with model without 2-way interaction
#modLMER2 <- lmer(observedLength ~ (factor(species) + emerge_detect_sumTScaled + emerge_detect_sumFScaled + ydayScaled +count_Scaled)^2 + 1|year, data = d_WB_BKT_BNT)
```

```{r calcrelipmm}
#https://gist.github.com/BERENZ/e9b581a4b7160357934e
calc.relip.mm <- function(model,type='lmg') {
  if (!isLMM(model) & !isGLMM(model)) {
    stop('Currently supports only lmer/glmer objects', call. = FALSE)
  }
  require(lme4)
  X <- getME(model,'X')
  X <- X[,-1]
  Y <- getME(model,'y')
  s_resid <- sigma(model)
  s_effect <- getME(model,'theta')*s_resid
  s2 <- sum(s_resid^2,s_effect^2)
  V <- Diagonal(x = s2,n=nrow(X))
  YX <- cbind(Y,X)
  cov_XY <- solve( t(YX) %*% solve(V) %*% as.matrix(YX))
  colnames(cov_XY) <- rownames(cov_XY) <- colnames(YX)
  importances <- calc.relimp(as.matrix(cov_XY),rela=T,type=type)
  return(importances)
}
```


```{r lmer, cache = TRUE}

#modLMER1 <- lmer(observedLength ~ (factor(species) + get(tTime)|year + get(fTime)|year + ydayScaled|year + count_Scaled|year), data = d_WB_BKT_BNT)

modLMER1 <- lmer(observedLength ~ (factor(species) * get(tTime) * get(fTime) * ydayScaled * count_Scaled) + (1|year), data = d_WB_BKT_BNT)

modLMER2 <- lmer(observedLength ~ (factor(species) + get(tTime) + get(fTime) + ydayScaled + count_Scaled) + (1|year), data = d_WB_BKT_BNT)

modLMER3 <- lmer(observedLength ~ (factor(species) + get(tTime) + get(fTime) + ydayScaled + count_Scaled)^2 + (1|year), data = d_WB_BKT_BNT)

# one-by-one
modLMER1a <- lmer(observedLength ~ (factor(species))  + (1|year), data = d_WB_BKT_BNT)
modLMER1b <- lmer(observedLength ~ (factor(species) * get(tTime))  + (1|year), data = d_WB_BKT_BNT)
modLMER1c <- lmer(observedLength ~ (factor(species) * get(fTime)) + (1|year), data = d_WB_BKT_BNT)
modLMER1d <- lmer(observedLength ~ (factor(species) * ydayScaled) + (1|year), data = d_WB_BKT_BNT)
modLMER1e <- lmer(observedLength ~ (factor(species) * count_Scaled) + (1|year), data = d_WB_BKT_BNT)

AIC(modLMER1, modLMER2, modLMER3, modLMER1a, modLMER1b, modLMER1c, modLMER1d, modLMER1e) %>% arrange(AIC)

calc.relip.mm(modLMER3)
library(MuMIn)
r.squaredGLMM(modLMER1)

summary(modLMER1)
ranef(modLMER1)
```



Raw data exploration following the models  

relaimpo::calc.relimp(modLM2) - other models are too big
                                 lmg
factor(species)          0.005339423
emerge_detect_sumTScaled 0.006883496
emerge_detect_sumFScaled 0.063450197
ydayScaled               0.012116064
count_Scaled             0.028031531
```{r plotRawInd}

ggplot(d_WB_BKT_BNT, aes(get(fTime), observedLength)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~species)

ggplot(d_WB_BKT_BNT, aes(count_Scaled, observedLength)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~species)

ggplot(d_WB_BKT_BNT, aes(ydayScaled, observedLength)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~species)

ggplot(d_WB_BKT_BNT, aes(get(tTime), observedLength)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~species)

```
